E Coli Project Paper Incredibly Rough Draft

### 2013-07-29 paper/paper.tex
took figures from paper

To test our computational model, we implemented a pill shaped cell, and
tested using the same cell parameters as Huang et al, which were a radius
of 0.5 $\mu$m in the middle and at the spherical endcaps, and two different
cell lengths of 4 $\mu$m and 10 $\mu$m. We found the same type of
oscillations as in their paper using these initial conditions, verifying
that our model works as intended. Below are snapshots of MinD and MinE
concentrations at 5 second timestamps in the 4 $\mu$m cell:
\newline
\newline
[insert 5 second time stamps of 4 $\mu$m sim]
\newline
\newline
We then began to define other, non-traditional cell shapes for the purpose of
modeling squished and perturbed E. coli cells, which were created
experimentally in Mannick et al. To achieve this, we went with a
cartesian lattice rather than the cylindrical lattice used in Huang et 
al's simulations, as it allows for more flexibility in defining the 
cell shape. Some of the cell shape models included a flattened pill
(stadium shape), an ellipsoid, a spherical cell, and various randomly
generated smooth shapes, such as those in the figures below. 
\newline
\newline
[insert memf print of 2-3 cell shapes]
\newline
\newline
To interpret the results, we generated several different plot views of
the printed simulation data. These plots included a time averaged view
of the protein densities in the cell; a plot tracking the location of
protein concentrations that were global maxima in space and local maxima in 
time; and an animated view that showed the actual dispersion of
protein concentrations in the cell over time. 
\subsection{Mathematical Model}
\section{Specific Results}
\subsection{Pill Shape}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/frequency_plot_all_norm-p-40-20-0-0-150.pdf}
  \caption{Protein density fluctuation with time against the polar
    wall in a 4$\mu m$ by 2$\mu m$ pill shaped cell.  The wall-bound
    proteins are shown at a point that is on the wall and the
    cytoplasm densities are shown at a point immediately adjecent.
    This plot is normalized so that every protein density peaks at 1}
  \label{frequency-plot-40-20-0-0-150}
\end{figure}

\fixme{Be more exact about what exactly Nd is versus nATP, in terms of
  the unit and dimensions.  The plotting may need to be changed}
Figure \ref{frequency-plot-40-20-0-0-150} shows the protein density
fluctuations at a point in space adjecent to a polar wall.  At each
collection of peaks the proteins reach their zenith in an order that
agrees with the qualitative picture described by
Huang\cite{huang2003dynamic}, except for nADP peak.  The peaks start
with a maxima of ATP-MinD accumulating at the walls \fixme{Need a
  better way to refer to compounds}. This peak is followed by a peak
in Nde, or ATP-MinD-MinE, as the cell is converted on the wall from
the former to the latter compound.  As this compound splits apart and
leaves the wall we see a peak in nE, or MinE in the cytoplasm.  This
is then followed by a broader peak in nATP, or the MinD-ATP compound,
as minD-ADP is (changed but I forgot the name for it) into minD-ATP in
the cytoplasm.  The minD-ATP naturally diffuses away from the pole,
which is shown by the broad nature of this peak.  All of these peaks
fit the qualitative picture except for the sharp MinD-ADP peak.  One
could except that it is a sharp peak, meaning that the MinD-ADP proteins
do not last in the cytoplasm for very long before they are converted
(once again forgot the name) into MinD-ATP.  The location of the peak
in the time dimension is troubling, however, since that we would
expect it to occur just after the peak in Nde, since the wall-bound
MinD-ATP-MinE proteins are the ones that split up and leave the wall,
creating both the MinE and the MinD-ADP.  One would expect to see the
minD-ADP peak to coincide in time with the minE peak, but it instead
coincides with the wall-bound MinD-ATP-MinE peak. (Although you can
maybe sort of convince yourself otherwiase, really looking at it).

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/frequency_plot_all_middle-p-40-20-0-0-150.pdf}
  \caption{A comparison of protein density fluctuation with time
    against the polar wall and at a point in the center of the
    z-length and adjacent to the wall, in the cytoplasm.  The cell is
    pill shaped and has dimensions of 4$\mu m$ by 2$\mu m$.  }
  \label{frequency-plot-40-20-0-0-150}
\end{figure}

The theory states that the aversion of the FtsZ polymer to the MinC
protein is responsible for the FtsZ polymer setting up in the center
of the cell as opposed to at the cellular poles.  The MinC protein in
turn natrually associates with the minD protein that is modelled in
Huang's differential equations.  \fixme{look this up and write this
  again} The theory is dependent on there being a significant
difference in protein density between the center and polar regions. It
is worthwhile examining this difference in detail.  Figure
\ref{frequency-plot-40-20-0-0-150} shows a comparison of density
fluctuation over time at a point adjacent to the polar cell wall and
at a point adjacent to the cell wall in the center region of the cell.
The different proteins show different relationships between their
maxima and minima.  The MinD-ADP show sharp spikes at the poles and
much smaller oscillations at in the middle.  The values at the poles
are always greater than those in the middle, and the polar peaks
exhibit densities that are roughly 2.5 times greater than the maxima
in the center (\fixme{this factor and the following factors will be put
  into a table of comparisons for each protein and for the different
  cell sizes}).  The MinD-ATP densities show a very different trend.
For these it is the center density that is almost always greater than
the polar density.  The difference in density at the pole versus the
middle is a factor of roughly .85 (once again we'll do this more
officially in a table).  The MinE is similar to the MinD-ADP
comparison.  It shows a difference in maxima by a factor of roughly 2.
The MinC protein is associated with the minD protein, so perhaps the
most important camparison is between the nfl-MinD (MinD protein in all
its forms) at the ends and in the center of the pill.  This shows
sharp spikes at the pills and a maxima difference factor of roughly
1.25. \fixme{I'd like to analyze further what this means for the FtZ
  polymer.  Would a difference like this really have a big effect?
  What sort of effect should it have, looking at experimental data?
  One thing is that Mannik (is it Mannik?) genetically deletes the
  MinC protein and they get 73 percent proper cell division.  It
  really would be good to do a more extensive analysis here,
  referencing what other papers say about this.}

@@

\section{animate?}
\animategraphics[height=2.8in,autoplay,controls]{12}{../data/shape-randst/plots/movie-density-Nd-96/animate_}{0}{120}

>>>>

To test our computational model, we implemented a pill shaped cell, and
tested using the same cell parameters as Huang et al, which were a radius
of 0.5 $\mu$m in the middle and at the spherical endcaps, and two different
cell lengths of 4 $\mu$m and 10 $\mu$m. We found the same type of
oscillations as in their paper using these initial conditions, verifying
that our model works as intended. Below are snapshots of MinD and MinE
concentrations at 5 second timestamps in the 4 $\mu$m cell:
\newline
\newline
[insert 5 second time stamps of 4 $\mu$m sim]
\newline
\newline
We then began to define other, non-traditional cell shapes for the purpose of
modeling squished and perturbed E. coli cells, which were created
experimentally in Mannick et al. To achieve this, we went with a
cartesian lattice rather than the cylindrical lattice used in Huang et
al's simulations, as it allows for more flexibility in defining the
cell shape. Some of the cell shape models included a flattened pill
(stadium shape), an ellipsoid, a spherical cell, and various randomly
generated smooth shapes, such as those in the figures below.
\newline
\newline
[insert memf print of 2-3 cell shapes]
\newline
\newline
To interpret the results, we generated several different plot views of
the printed simulation data. These plots included a time averaged view
of the protein densities in the cell; a plot tracking the location of
protein concentrations that were global maxima in space and local maxima in
time; and an animated view that showed the actual dispersion of
protein concentrations in the cell over time.
\subsection{Mathematical Model}
\section{Specific Results}
\subsection{Pill Shape}
%% \begin{figure}
%%   \includegraphics[width=\columnwidth]{../data/shape-p/plots/frequency-plot-all-NflD-p-40-20-0-0-150.pdf}
%%   \caption{Protein density fluctuation with time against the polar
%%     wall in a 4$\mu m$ by 2$\mu m$ pill shaped cell.  The wall-bound
%%     proteins are shown at a point that is on the wall and the
%%     cytoplasm densities are shown at a point immediately adjecent.
%%     This plot is normalized so that every protein density peaks at 1}
%%   \label{frequency-plot-40-20-0-0-150}
%% \end{figure}

\fixme{Be more exact about what exactly Nd is versus nATP, in terms of
  the unit and dimensions.  The plotting may need to be changed}
Figure %%\ref{frequency-plot-40-20-0-0-150} shows the protein density
fluctuations at a point in space adjecent to a polar wall.  At each
collection of peaks the proteins reach their zenith in an order that
agrees with the qualitative picture described by
Huang\cite{huang2003dynamic}, except for nADP peak.  The peaks start
with a maxima of ATP-MinD accumulating at the walls \fixme{Need a
  better way to refer to compounds}. This peak is followed by a peak
in Nde, or ATP-MinD-MinE, as the cell is converted on the wall from
the former to the latter compound.  As this compound splits apart and
leaves the wall we see a peak in nE, or MinE in the cytoplasm.  This
is then followed by a broader peak in nATP, or the MinD-ATP compound,
as minD-ADP is (changed but I forgot the name for it) into minD-ATP in
the cytoplasm.  The minD-ATP naturally diffuses away from the pole,
which is shown by the broad nature of this peak.  All of these peaks
fit the qualitative picture except for the sharp MinD-ADP peak.  One
could except that it is a sharp peak, meaning that the MinD-ADP proteins
do not last in the cytoplasm for very long before they are converted
(once again forgot the name) into MinD-ATP.  The location of the peak
in the time dimension is troubling, however, since that we would
expect it to occur just after the peak in Nde, since the wall-bound
MinD-ATP-MinE proteins are the ones that split up and leave the wall,
creating both the MinE and the MinD-ADP.  One would expect to see the
minD-ADP peak to coincide in time with the minE peak, but it instead
coincides with the wall-bound MinD-ATP-MinE peak. (Although you can
maybe sort of convince yourself otherwiase, really looking at it).

%% \begin{figure}
%%   \includegraphics[width=\columnwidth]{../data/shape-p/plots/frequency_plot_all_middle-p-40-20-0-0-150.pdf}
%%   \caption{A comparison of protein density fluctuation with time
%%     against the polar wall and at a point in the center of the
%%     z-length and adjacent to the wall, in the cytoplasm.  The cell is
%%     pill shaped and has dimensions of 4$\mu m$ by 2$\mu m$.  }
%%   \label{frequency-plot-40-20-0-0-150}
%% \end{figure}

The theory states that the aversion of the FtsZ polymer to the MinC
protein is responsible for the FtsZ polymer setting up in the center
of the cell as opposed to at the cellular poles.  The MinC protein in
turn natrually associates with the minD protein that is modelled in
Huang's differential equations.  \fixme{look this up and write this
  again} The theory is dependent on there being a significant
difference in protein density between the center and polar regions. It
is worthwhile examining this difference in detail.  Figure
%%\ref{frequency-plot-40-20-0-0-150} shows a comparison of density
fluctuation over time at a point adjacent to the polar cell wall and
at a point adjacent to the cell wall in the center region of the cell.
The different proteins show different relationships between their
maxima and minima.  The MinD-ADP show sharp spikes at the poles and
much smaller oscillations at in the middle.  The values at the poles
are always greater than those in the middle, and the polar peaks
exhibit densities that are roughly 2.5 times greater than the maxima
in the center (\fixme{this factor and the following factors will be put
  into a table of comparisons for each protein and for the different
  cell sizes}).  The MinD-ATP densities show a very different trend.
For these it is the center density that is almost always greater than
the polar density.  The difference in density at the pole versus the
middle is a factor of roughly .85 (once again we'll do this more
officially in a table).  The MinE is similar to the MinD-ADP
comparison.  It shows a difference in maxima by a factor of roughly 2.
The MinC protein is associated with the minD protein, so perhaps the
most important camparison is between the nfl-MinD (MinD protein in all
its forms) at the ends and in the center of the pill.  This shows
sharp spikes at the pills and a maxima difference factor of roughly
1.25. \fixme{I'd like to analyze further what this means for the FtZ
  polymer.  Would a difference like this really have a big effect?
  What sort of effect should it have, looking at experimental data?
  One thing is that Mannik (is it Mannik?) genetically deletes the
  MinC protein and they get 73 percent proper cell division.  It
  really would be good to do a more extensive analysis here,
  referencing what other papers say about this.}

@@

\section{animate?}
%%\animategraphics[height=2.8in,autoplay,controls]{12}{../data/shape-randst/plots/movie-density-Nd-96/animate_}{0}{120}
